\documentclass[8pt]{beamer}
\usepackage[]{graphicx}
\usepackage[]{color}


\usetheme{metropolis}           % Use metropolis theme
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{tabularx}

\usepackage[style=authoryear]{biblatex}
\addbibresource{references.bib}
\usepackage{cleveref}
\renewcommand*{\bibfont}{\footnotesize}


\title{Budget Deep Dive}
\author{Ryan Giordano}
\date{Sep 23rd, 2022}
\institute{Children's Community Center}

\begin{document}




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Outline}

Here are some questions you might have:
%
\begin{itemize}
%
\item Ryan, do we have money for this thing I want to do?
\item How is that that we have so much money in the bank, especially after COVID?
\item If we have so much money in the bank, why didn't we give the teachers
    a raise?
\item How does the budget get set anyway, and how good are we at meeting it?
\item How will equity tuition change all of this?
%
\end{itemize}
%
I will try to sketch answers to these questions today.
%
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Who are the budgeteers?}
%
\begin{itemize}
%
\item Vidrik Frankfather, our staff accountant.
\begin{itemize}
    \item A past CCC parent and board treasurer
    \item Has actual accounting expertise
    \item Discovered errors in the work done by our contracted accounting firm,
        suggested that he could do it better as a part-time staff member
    \item A great repository of detailed information and historical context
\end{itemize}
%
%
\item Edna Tow, our school manager.
\begin{itemize}
    \item Also a past CCC parent
    \item Collects tuition, pays salaries, processes reimbursements
    \item Manages bank accounts and investments
    \item A great repository of detailed information and historical context
\end{itemize}
%
%
\item Board treasurer (whomever they may be)
\begin{itemize}
    \item May not have actual accounting experience
    \item May not know a lot of historical context
    \item May not have even been present when the last budget was passed
    \item Facilitates communication between parents, the board, the teachers,
        and the staff who actually know what they are doing
\end{itemize}
%
\end{itemize}
%
\end{frame}



\end{document}
